---
title: Investigate the exponential distribution in R and compare it with the Central
  Limit Theorem
author: "Arnav Pant"
date: "3rd October 2024"
output:
  html_document: default
  pdf_document: default
---

## Overview
The purpose of this analysis is to understand exponential distribution and compare it to the Central Limit Theorem.
We'll set lambda = 0.2, simulate 1000 averages of 40 exponentials, and perform the following tasks:

       1) Sample Mean vs Theoretical Mean
       2) Sample Variance vs Theoretical Variance
       3) Distribution Approximation to Normal


## Simulations

Set the simulation variables lambda, exponentials, and seed.
```{r}
ECHO=TRUE
set.seed(1337)
lambda = 0.2
exponentials = 40
```

Run Simulations with variables
```{r}
simMeans = NULL
for (i in 1 : 1000) simMeans = c(simMeans, mean(rexp(exponentials, lambda)))
```


## Sample Mean versus Theoretical Mean

#### Sample Mean
Calculating the mean from the simulations with give the sample mean.
```{r}
mean(simMeans)
```

#### Theoretical Mean
The theoretical mean of an exponential distribution is lambda^-1.
```{r}
lambda^-1
```

#### Comparison
There is a difference in the simulations sample mean and the exponential distribution's theoretical mean.
```{r}
abs(mean(simMeans)-lambda^-1)
```


## Sample Variance versus Theoretical Variance

#### Sample Variance
Calculating the variance from the simulation means with give the sample variance.
```{r}
var(simMeans)
```

#### Theoretical Variance
The theoretical variance of an exponential distribution is (lambda * sqrt(n))^-2.
```{r}
(lambda * sqrt(exponentials))^-2
```

#### Comparison
There is a slight difference between the simulations sample variance and the exponential distribution's theoretical variance.
```{r}
abs(var(simMeans)-(lambda * sqrt(exponentials))^-2)
```


## Distribution
This is a density histogram of the 1000 simulations. There is an overlay with a normal distribution that has a mean of lambda^-1 and standard deviation of (lambda*sqrt(n))^-1, the theoretical normal distribution for the simulations.
```{r}
library(ggplot2)
ggplot(data.frame(y=simMeans), aes(x=y)) + 
  geom_histogram(aes(y=..density..), binwidth=0.2, fill="#0072B2", 
                 color="black") +
  stat_function(fun=dnorm, arg=list(mean=lambda^-1, 
                                    sd=(lambda*sqrt(exponentials))^-1), 
                size=2) +
  labs(title="Plot of the Simulations", x="Simulation Mean")
```